---
title: "Project 3 EDA"
author: "Susan Hajmohammad"
format: pdf
---

Introduction section
You should have an introduction section that
• briefly describes the data and the variables you have to work with.
• describes the purpose of your EDA and ultimate goal of modeling.

## Introduction: 

Diabetes is a chronic illness that affects more than 36 million Americans and many more people world-wide.  Diabetes is a condition where the body’s cells don't respond to insulin properly, which leads to high blood sugar levels. It is currently understood to be irreversible and can result in a number of life altering comorbidities.  Studies shows that most cases of type 2 diabetes can be prevented through lifestyle changes.  As such, it is important to explore how certain health predictors are at predicting an outcome of diabetes because that would be a useful tool to predicting patient outcomes and guiding lifestyle interventions.  

With a lot of research already having been done on this topic, certain health measurements are more obviously correlated with outcomes of diabetes (BMI, cholesterol, eating habits); so, for this EDA I wanted to explore some less obvious predictors of diabetes.  I will be looking at Age, Sex, Education level and Income level as predictors of Diabetes diagnosis.  


## Libraries: 

```{r}
#| warning: FALSE
#| message: FALSE

library(tidyverse)
library(ggplot2)

```


## Data: 

Import data and change levels to meaningful names. 


```{r}
#read in diabetes data from folder
diabetes_data <- read_csv("diabetes_data.csv")

#now for chosen variables change each to factor and give meaningful labels for the levels
diabetes_data_reduced <- read_csv("diabetes_data.csv")%>% 
  mutate(Diabetes_binary = factor(Diabetes_binary, levels= c(0,1), labels=c("No Diabetes", "Diabetes")),
         Sex = factor(Sex, levels= c(0,1), labels = c("Female", "Male")), 
         # info sheet says 5 year increments
         Age = factor(Age, levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13), labels = c("18to24", "25to29", "30to34", "35to39", "40to44", "45to49", "50to54", "55to59", "60to64", "65to69", "70to74", "75to79", "80plus")),
         Education= factor(Education, levels= c(1,2,3,4,5,6), labels=c("Never attended or only kindergarten","Elementary","Some high school", "High school graduate","Some college or technical school","College graduate")), 
         Income = factor(Income, levels = c(1,2,3,4,5,6,7,8), labels = c("< $10,000","< $15,000","< $20,000","< $25,000", "< $35,000","< $50,000","< $75,000","$75,000+")))%>%
  #select only relevant variables to work with smaller data set
  select(Diabetes_binary, Age, Sex, Education, Income)

```

## Univariate analyses: 

### Histograms for each variable: 

```{r}

#age 

a <- ggplot(diabetes_data_reduced, aes(x=Age))
a + geom_bar()

#sex

s <- ggplot(diabetes_data_reduced, aes(x=Sex))
s+ geom_bar()

#Education
e <- ggplot(diabetes_data_reduced, aes(x=Education))
e+ geom_bar()

#Income
i <- ggplot(diabetes_data_reduced, aes(x=Income))
i+ geom_bar()

```












[Click here for the Modeling Page](Modeling.html)
